---
title: "Compare wind products"
author: "EE Holmes"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source(file.path(here::here(), "inst", "extdata", "get_satellite_data", "Replicate_EMT", "get_EMT_functions.R"))
library(ggplot2)
library(tidyr)
library(dplyr)
library(raster)
library(rasterVis)
library(plotrix)
```

# Comparison of daily winds across products

Compare the wind estimates to that reported in other sources.

```{r}
dates <- c("2010-07-28", "2010-07-30")
lats <- c(6, 20); lons <- c(68,72)
wind1 <- getdata("erdlasFnPres6", date=dates, lat=lats, lon=lons)
wind1 <- getwindfromP(wind1)
wind1$date <- as.Date(wind1$time)
wind1.day <- wind1 %>% 
  group_by(date, latitude, longitude) %>% 
  summarize(u = mean(u, na.rm=TRUE), v=mean(v, na.rm=TRUE))
```

Now get ASCAT wind. This is on a finer grid. 
```{r}
wind2 <- getdata("erdQAwind1day", pars=c("x_wind", "y_wind"), date=dates, lat=lats, lon=lons)
wind2$date <- as.Date(wind2$time)
res <- attr(wind2, "resolution")
```

```{r}
wind2.day <- wind2[,c("date", "latitude", "longitude", "x_wind", "y_wind")]
colnames(wind2.day) <- c("date", "latitude", "longitude", "u", "v")
```

Now get CCMP 6-hourly data and average to daily.
```{r}
wind3 <- getdata("jplCcmpL3Wind6Hourly", date=dates, lat=lats, lon=lons)
wind3$date <- as.Date(wind3$time)
wind3.day <- wind3 %>% 
  group_by(date, latitude, longitude) %>% 
  summarize(u = mean(uwnd, na.rm=TRUE), v=mean(vwnd, na.rm=TRUE))
wind3.day$longitude <- wind3.day$longitude-0.125
wind3.day$latitude <- wind3.day$latitude-0.125
```

### Plot data

```{r}
df <- rbind(data.frame(wind1.day, data="FNMOC"),
            data.frame(wind2.day, data="ASCAT"),
            data.frame(wind3.day, data="CCMP"))
dfl <- df %>%
  pivot_longer(!date & !latitude & !longitude & !data,
                     names_to = "name",
                     values_to = "wind")
lon <- 69.5; thedate <- max(dfl$date)
p1 <- ggplot(subset(dfl, longitude==lon & date==thedate), aes(x=latitude, y=wind, color=data)) + geom_line() +
  facet_wrap(~name) + ggtitle(paste(thedate, "longitude =", lon))
lon <- 71.5; thedate <- min(dfl$date)
p2 <- ggplot(subset(dfl, longitude==lon & date==thedate), aes(x=latitude, y=wind, color=data)) + geom_line() +
  facet_wrap(~name) + ggtitle(paste(thedate, "longitude =", lon))
gridExtra::grid.arrange(p1,p2)
```

Plot the u and v across latitudes.
```{r}
p1 <- ggplot(subset(df, date==thedate), aes(x=latitude, y=u, color=as.factor(longitude))) + geom_line() + facet_wrap(~data) + theme(legend.position = "none")
p2 <- ggplot(subset(df, date==thedate), aes(x=latitude, y=v, color=as.factor(longitude))) + geom_line() + facet_wrap(~data) + theme(legend.position = "none")
gridExtra::grid.arrange(p1,p2)
```

# Test 2: Comparison of monthly winds 2010

Compare the wind estimates to that reported in other sources.

```{r}
dates <- c("2010-01-01", "2011-01-01")
lats <- c(6, 20); lons <- c(68,72)
windmon1 <- getdata("erdlasFnWPr", date=dates, lat=lats, lon=lons)
windmon1$date <- format(windmon1$time, "%Y-%m-01")
colnames(windmon1) <- stringr::str_replace(colnames(windmon1), "_mean", "")
windmon1 <- getEMT(windmon1, coast_angle=158)
```

Now get ASCAT monthly wind. This is on a finer grid. 
```{r}
#10m winds
tmp <- getdata("erdQAwindmday", date=dates, lat=lats, lon=lons, altitude=10)
windmon2 <- getdata("erdQAstressmday", date=dates, lat=lats, lon=lons, altitude=0)
windmon2$u <- tmp$x_wind
windmon2$v <- tmp$y_wind
windmon2$uv_mag <- sqrt(tmp$y_wind^2 + tmp$x_wind^2)
windmon2$date <- format(windmon2$time, "%Y-%m-01")
res <- attr(windmon2, "resolution")
windmon2 <- getEMT(windmon2, coast_angle=158)
```

Now get CCMP monthly.
```{r}
windmon3 <- getdata("jplCcmp35aWindMonthly", date=dates, lat=lats, lon=lons)
windmon3$date <- format(windmon3$time, "%Y-%m-01")
colnames(windmon3) <- stringr::str_replace(colnames(windmon3), "wnd", "")
colnames(windmon3) <- stringr::str_replace(colnames(windmon3), "wspd", "uv_mag")
windmon3$longitude <- windmon3$longitude-0.125
windmon3$latitude <- windmon3$latitude-0.125
windmon3 <- getEMT(windmon3, coast_angle=158)
```

Now get the Reanalysis Data ERA5 monthly wind product. 
```{r}
# the altitude is in millibar. air pressue at 10m is about 1000millibar
#h=10 #meters
#airmillibar <- 0.01*101325*(1 - 2.25577*10^-5*h)^5.25588 #pascal/100

windmon4 <- getdata("hawaii_soest_66d3_10d8_0f3c", date=dates, lat=lats, lon=lons, 
                    altitude=1000, eserver="http://apdrc.soest.hawaii.edu/erddap", alt.name="LEV")
windmon4$date <- format(windmon4$time, "%Y-%m-01")
windmon4 <- getEMT(windmon4, coast_angle=158)
```

### Plot wind data

```{r}
library(tidyr)
cols <- c("date", "latitude", "longitude", "u", "v", "taux","tauy", "ektrx", "ektry", "EMTperp")
df <- rbind(data.frame(windmon1[,cols], data="FNMOC"),
            data.frame(windmon2[,cols], data="ASCAT"),
            data.frame(windmon3[,cols], data="CCMP"),
            data.frame(windmon4[,cols], data="ERA5"))
dfl <- df %>%
  pivot_longer(!date & !latitude & !longitude & !data,
                     names_to = "name",
                     values_to = "value")
```

Winds.
```{r}
library(ggplot2)
thedate <- "2010-07-01"
plotlons <- c(69.5, 71.5)
pars <- c("u", "v")
p1 <- ggplot(subset(dfl, longitude==plotlons[1] & date==thedate & name%in%pars), aes(x=latitude, y=value, color=data)) + geom_line() +
  facet_wrap(~name) + ggtitle(paste(thedate, "longitude =", plotlons[1]))
p2 <- ggplot(subset(dfl, longitude==plotlons[2] & date==thedate & name%in%pars), aes(x=latitude, y=value, color=data)) + geom_line() +
  facet_wrap(~name) + ggtitle(paste(thedate, "longitude =", plotlons[2]))
gridExtra::grid.arrange(p1,p2)
```



Wind stress
```{r}
thedate <- "2010-07-01"
plotlons <- c(69.5, 71.5)
pars <- c("taux", "tauy")
p1 <- ggplot(subset(dfl, longitude==plotlons[1] & date==thedate & name%in%pars), aes(x=latitude, y=value, color=data)) + geom_line() +
  facet_wrap(~name) + ggtitle(paste(thedate, "longitude =", plotlons[1]))
p2 <- ggplot(subset(dfl, longitude==plotlons[2] & date==thedate & name%in%pars), aes(x=latitude, y=value, color=data)) + geom_line() +
  facet_wrap(~name) + ggtitle(paste(thedate, "longitude =", plotlons[2]))
gridExtra::grid.arrange(p1,p2)
```

Ekman Mass Transport

```{r emt.plot}
thedate <- "2010-07-01"
plotlons <- c(69.5, 71.5)
pars <- c("ektrx", "ektry", "EMTperp")
p1 <- ggplot(subset(dfl, longitude==plotlons[1] & date==thedate & name%in%pars), aes(x=latitude, y=value, color=data)) + geom_line() +
  facet_wrap(~name, scales = "free_y") + ggtitle(paste(thedate, "longitude =", plotlons[1]))
p2 <- ggplot(subset(dfl, longitude==plotlons[2] & date==thedate & name%in%pars), aes(x=latitude, y=value, color=data)) + geom_line() +
  facet_wrap(~name, scales = "free_y") + ggtitle(paste(thedate, "longitude =", plotlons[2]))
gridExtra::grid.arrange(p1,p2)
```

# Test 3: Comparison of EMT perpendicular using monthly winds 2000 to 2019


```{r}
b=as(as(indiashp2, "SpatialLinesDataFrame"), "SpatialPointsDataFrame")
b=subset(b, Line.NR==129)
coastcoord=coordinates(b)
coastcoord <- coastcoord[coastcoord[,"x"]<77.5,]
coastlats = unique(windmon3.mean$latitude)
coastlons = c()
for(i in coastlats){
  tmp <- coastcoord[,"y"]
  coastlons <- c(coastlons, min(coastcoord[,"x"][which(abs(tmp-i)==min(abs(tmp-i)))]))
}
```

```{r}
dates <- c("1999-08-16", "2009-10-16")
lats <- c(8, 13); lons <- c(70,78)
```

```{r windmon3.coast, echo=FALSE, warning=FALSE}
windmon1.coast <- windmon1[windmon3$latitude>7,]
for(i in 1:length(coastlats)){
  thelat <- coastlats[i]
  thelon <- coastlons[i]
  windmon3.coast <- windmon3.coast[!(windmon3.coast$latitude==thelat & windmon3.coast$longitude<(thelon-2)),]
}
windmon3.cmean <- windmon3.coast %>% 
  group_by(month, latitude, longitude) %>% 
  summarize(ektrx = mean(ektrx, na.rm=TRUE), 
            ektry = mean(ektry, na.rm=TRUE),
            upi = mean(upi, na.rm=TRUE),
            v = mean(v, na.rm=TRUE),
            u = mean(u, na.rm=TRUE),
            EMT = mean(EMT, na.rm=TRUE),
            EMTperp = mean(EMTperp, na.rm=TRUE)
)
```

```{r windmon3.coast.plot, echo=FALSE, warning=FALSE}
for(i in month.name){
rdf <- subset(windmon3.cmean, month==i)
rdf$x <- rdf$longitude
rdf$y <- rdf$latitude
rdf <- rdf[, c("x", "y", "EMTperp")]
if(i=="January"){ rst <- rasterFromXYZ(rdf); names(rst) <- i }else{
  tmp <- rasterFromXYZ(rdf); names(tmp) <- i
  rst <- stack(rst, tmp)
}
}
at=seq(-3000,2000,100)
pal <- colorRampPalette(c("red","white","blue"))
spplot(rst, col.regions=rev(pal(length(at))), at=at, zlim=c(-3000,500),
       main="Average Monthly EMT perp to coast (+ in, - out)") + layer(sp.polygons(indiashp2))
```


```{r windmon3.coast.EMTx.plot, echo=FALSE, warning=FALSE}
for(i in month.name){
rdf <- subset(windmon3.cmean, month==i)
rdf$x <- rdf$longitude
rdf$y <- rdf$latitude
rdf <- rdf[, c("x", "y", "ektrx")]
if(i=="January"){ rstew <- rasterFromXYZ(rdf); names(rstew) <- i }else{
  tmp <- rasterFromXYZ(rdf); names(tmp) <- i
  rstew <- stack(rstew, tmp)
}
}
at=seq(-3000,2000,100)
pal <- colorRampPalette(c("red","white","blue"))
spplot(rstew, col.regions=rev(pal(length(at))), at=at, zlim=c(-3000,500),
       main="Average Monthly EMT East(+) West(-)") + layer(sp.polygons(indiashp2))
```


Mean Ekman Transport perpendicular to coast line.

```{r echo=FALSE}
plot(tapply(-1*windmon3.cmean$EMTperp, windmon3.cmean$month, mean, na.rm=TRUE)[month.name], type="l",
     main="Mean Ekman Transport Perp to Coast 1999-2017 (+ = offshore)", ylab="-1 x EMT perp (kg/ms)", xlab="month",
     xaxt="n")
axis(1, at=1:12, labels=month.abb)
```


Mean Ekman Transport east-west.

```{r echo=FALSE}
plot(tapply(-1*windmon3.cmean$ektrx, windmon3.cmean$month, mean, na.rm=TRUE)[month.name], type="l",
     main="Mean Ekman Transport East-West 1999-2017 (+ = offshore)", ylab="-1 x EMT perp (kg/ms)", xlab="month",
     xaxt="n")
axis(1, at=1:12, labels=month.abb)
```

